---
title: "stl_lead_inequity"
author: "Felicia Cruz"
date: "8/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
```

## Read in the data 
```{r}
stl_lead <- read_delim(here::here("data", "stl_blood_lead.csv"), delim = ";")
```

## Data exploration
```{r}
names(stl_lead)
stl_lead <- stl_lead %>% clean_names()
```


```{r}
stl_lead_prop <- stl_lead %>% mutate(prop_white = (white / total_pop) * 100)
```

## Create a scatterplot 
```{r stl_lead_plot, fig.height = 4, fig_width = 6}
stl_lead_prop <- stl_lead_prop %>% 
  mutate(pct_elevated = str_replace(string = pct_elevated, pattern = ",", replacement = "."))
# fixes the decimal places in pct_elevated

stl_lead_plot <- ggplot(data = stl_lead_prop, aes(x = prop_white, y = as.numeric(pct_elevated))) + 
  geom_point(aes(size = prop_white), 
             show.legend = FALSE,
             color = "magenta4",
             alpha = 0.5) + 
  labs(title = "High Blood Level Tests and Race by Census Tract", 
       x = "% White",
       y = "% Elevated")

stl_lead_plot
```

## save as a png to figs
```{r}
ggsave(here::here("figs", "stl_lead_plot.png"), plot = stl_lead_plot, width = 6, height = 5)
```
There is a negative correlation between the proportion of white children in a census tract and the percentage of elevated blood lead levels. As the percentage of white people increases in each census tract, the proportion of elevated lead concentration in blood decreases. 


## Create a histogram 
```{r}
hist <- ggplot(data = stl_lead_prop, aes(x = as.numeric(pct_elevated))) +
  geom_histogram(color = "cyan4",
                 fill = "magenta4",
                 size = 1) + 
  labs(title = "Elevated Blood Lead Levels",
       x = "% Elevated")

hist

```
## Export as a jpg to the figs folder 
```{r}
ggsave(here::here("figs", "lead_hist.jpg"), plot = hist, width = 6, height = 5)
```

